---
title: "Hypergeometric Monte Carlo"
# output: html_notebook
output: rmarkdown::github_document
---
This is a Monte Carlo sampling of the hypergeometric distribution


First load the libraries we will use
```{r, message=FALSE, warning=FALSE}
library(clusterSim)
library(ggplot2)
library(reshape2)
```
Information on R's hypergeometric distribution function can be found using this command
```{r}
#?Hypergeometric
```
A function to Monte Carlo sample from a hypergeometric distribution
```{r}
##  hg_mc - a function that monte carlo samples from a hypergeometric distribution
##  For each x in N, where N is the population size and x is the true number of successes in N, 
##  this function will draw a random sample of specified size (@sample) a specified number of times(@trials). 
##  @N - total population size
##  @samples - the number of samples drawn 
##  @trials - the number of times to perform independent random draws
##  @returns a dataframe witn N+1 columns (b/c 0 is included) and @samples+1 rows (O included)

hg_mc<-function(samples, N, trials){
  df <- NULL
  for (i in 0:N){
    df <- cbind(df, rhyper(trials, i, N-i, samples))
  }
  return(as.data.frame(df))
}
```
A helper function to aggregate the raw data from the simulations
```{r}
#helper function to aggregate successes 

sums <- function(samples, df){
  sums<- NULL
  for (i in 1:(ncol(df))){
    res <- c()
    for (j in 0:samples){
      res <- c(res, sum(df[ ,i] == j))
    }
    sums <- cbind(sums, res)
  }
  return(sums)
}
```
First we set the parameters we want. 
Here we set a population of 50, sample size of 6, and 10,000 trials. 
```{r}
## Set desired parameters for the simulation
samples <- 6  # sample size
N <- 50 # population size 
trials <- 10000 # number of trials
```
We now run the simulation and aggregate the raw data
```{r}
##  run the simulation
raw_data = hg_mc(samples, N, trials)

##  aggregate the data and rename the rows and cols
aggregated_data = sums(samples, raw_data)
colnames(aggregated_data) <- 0:N
rownames(aggregated_data) <- 0:samples
```
Let's look at the aggregated data. Each row is the observed number of successes in the sample.  Each column is the true number of successes in the population. 
```{r}
print(aggregated_data)
```
Let's create some plots to help visualize the data
First we write a function to plot the the distribution of the observed successes in a sample over the true number of successes in the population.
```{r}
##  plotting function
##  df is a dataframe of normalized aggregate successes from simulation
pdf_plot <- function(df, N){
  sample_size = ncol(df)-1
  test <- df
  test['successes'] <- 0:N
  melted = melt(test, id.vars="successes")
  ggplot() + 
    geom_line(data=melted, aes(x=successes, y=value, group=variable, color = variable), size=1) +
    labs(title=paste("Sample Size of", sample_size, " "), x =paste("True number of successes in population size ", toString(N), sep=" "), y = "Probability density") +
    guides(color=guide_legend(title="Successes\nin Sample"))
}
```
Call the plotting function 
```{r, message=FALSE, warning=FALSE}
##  normalize the results by successes in N to get a pdf
agg_data_normed <- data.Normalization(aggregated_data, type = "n10", normalization = "row")

##  plot the pdf's of the draw successes.  Data parameter is transposed and ensured to be a dataframe
pdf_plot(as.data.frame(t(agg_data_normed)), N)
```

This next plotting function is to examine the distribution of sample successes for a specific number of successes in the population.  Multiple population successes are allowed for comparison. 
```{r}
##  To plot the distribution of successes in samples for a specified number of successes in the population
##  @successes - the number of successes in the population
##  @df - a data frame containing the data
##  @N - the population size
 
draw_dist_plot <-function(successes, df, N){
  sample_size = nrow(df) -1
  df <- df[c(successes, 'count')]
  melted <- melt(df, id.vars='count')
  ggplot(data=melted, aes(x=count, y=value, fill=variable)) +
    geom_bar(stat="identity", position=position_dodge()) +
    labs(title=paste('Probability of successes in sample when true population successes are \n ', toString(successes), ' out of ', N), x=paste('Successes in sample size of ', sample_size), y='Probability Density') +
    stat_smooth(aes(color=variable), method='auto', se = FALSE)
}
```


We first set the population successes that we want to look at.  Here we choose 20 and 30.  
```{r, message=FALSE, warning=FALSE}
##  Plot entire sample distribution for successes in N.
successes=c(20, 30)
```
We normalize the data and call the plotting function
```{r, message=FALSE, warning=FALSE}
##  first normalize by sample to get pmf and add successes column
samples_normed <- as.data.frame(data.Normalization(aggregated_data, type = "n10", normalization = "col"))
samples_normed['count'] = c(0:(nrow(samples_normed)-1))
##  plot the distribution over successes in draws for specified successes in N
draw_dist_plot(successes, samples_normed, N)
```

If we want to examine a specific number of successes in a sample and find the likelihood of the true number of population successes we can plot the sample and sum the area under the curve. This function does that for the right tail of the distribution. 
```{r}
right_tail_plot <- function(df, sample_successes, pop_successes){
  sample_successes <- sample_successes +1
  df1 <- as.data.frame(t(df))
  df1['successes'] <- c(0:(nrow(df1)-1))
  shade <- df1[(pop_successes+1):nrow(df1),]
  prob <- 100* sum(shade[,sample_successes])
  ggplot(data=df1, aes(x=successes)) + geom_line(aes(y=df1[,sample_successes]), color='blue') + 
    geom_area(data=shade, aes(x=pop_successes:(nrow(df1)-1), y=shade[,sample_successes]), fill='blue') +
    labs(title=paste('Curve: ', sample_successes-1,' successes in sample of', toString(nrow(df)-1), '\nShaded Area: P[True Population Successes > ',
                     pop_successes, '] = ', prob, '%'), x='Successes in population', 
         y='Probability')
}
```
We then call the function based on the parameters we want to see.  Here we are looking at 4 successes in the sample size of 6 where we want to know the probability that the true number of successes in the population is greater than 25.  For a sample size or 6 with 4 successes, the probability that the true number of successes in the population of 50 is greater than 25 is roughly 80%
```{r}
##  parameters for plotting function
sample_successes <- 4 # number of successes in sample
pop_successes <- 25 # probability that true population successes are greater than this when observing sample_successes

right_tail_plot(agg_data_normed, sample_successes, pop_successes)
```

