---
title: "Hypergeometric Monte Carlo"
# output: html_notebook
output: rmarkdown::github_document
---
This is a Monte Carlo sampling of the hypergeometric distribution


First load the libraries we will use
```{r, message=FALSE, warning=FALSE}
library(clusterSim)
library(ggplot2)
library(reshape2)
```


Information on R's hypergeometric distribution function can be found using this command
```{r}
?Hypergeometric
```


A function to Monte Carlo sample from a hypergeometric distribution
```{r}
##  hg_sims simulates a draw of size @sample from a hypergeometric population of size @N @trials times
##  for each combination of successes and failures in @N (i.e. if @N = 10 then combinations are 
##  (0 successes, 10 failures), (1 successes, 9 failures), ..., (10 successes, 0 failures)).
##  @N - total population size
##  @samples - the number of samples drawn 
##  @trials - the number of times to perform independent random draws
##  @returns a dataframe witn N+1 columns (b/c 0 is included) and samples+1 rows (O included)

hg_sims<-function(samples, N, trials){
  df <- NULL
  for (i in 0:N){
    df <- cbind(df, rhyper(trials, i, N-i, samples))
  }
  return(as.data.frame(df))
}
```

A helper function to aggregate the raw data from the simulations
```{r}
#helper function to aggregate successes 

sums <- function(samples, df){
  sums<- NULL
  for (i in 1:(ncol(df))){
    res <- c()
    for (j in 0:samples){
      res <- c(res, sum(df[ ,i] == j))
    }
    sums <- cbind(sums, res)
  }
  return(sums)
}
```
Some plotting functions
```{r}
##  plotting function
##  df is a dataframe of normalized aggregate successes from simulation
pdf_plot <- function(df, N){
  sample_size = ncol(df)-1
  test <- df
  test['successes'] <- 0:N
  melted = melt(test, id.vars="successes")
  ggplot() + 
    geom_line(data=melted, aes(x=successes, y=value, group=variable, color = variable), size=1) +
    labs(title=paste("Sample Size of", sample_size, " "), x =paste("Successes out of ", toString(N), sep=" "), y = "Probability") +
    guides(color=guide_legend(title="Successes\nin Sample"))
}

##  To plot the distribution of successes in samples for a specified number of successes in the population
 
draw_dist_plot <-function(successes, df, N){
  sample_size = nrow(df) -1
  df <- df[c(successes, 'count')]
  melted <- melt(df, id.vars='count')
  ggplot(data=melted, aes(x=count, y=value, fill=variable)) +
    geom_bar(stat="identity", position=position_dodge()) +
    labs(title=paste('Probability of successes in sample when true population successes are \n ', toString(successes), ' out of ', N), x=paste('Successes in sample size of ', sample_size), y='Probability Density') +
    stat_smooth(aes(color=variable), method='auto', se = FALSE)
  
}
```

Run the simulations. Here we have a population of 50, sample size of 6, and 10,000 trials at each true number of successes out of 50. 
```{r}
## Set desired parameters for the simulation
samples <- 6  # sample size
N <- 50 # population size 
trials <- 50000 # number of trials

##  run the simulation
raw_data = hg_sims(samples, N, trials)

##  aggregate the data and rename the rows and cols
aggregated_data = sums(samples, raw_data)
colnames(aggregated_data) <- 0:N
rownames(aggregated_data) <- 0:samples
```

Look at the aggregated data. Each row is the observed number of successes.  Each column is the true number of successes in the population. 
```{r}
print(aggregated_data)
```




Plot of each sample success distribution over true successes in population
```{r, message=FALSE, warning=FALSE}
##  normalize the results by successes in N to get a pdf
agg_data_normed <- data.Normalization(aggregated_data, type = "n10", normalization = "row")

##  plot the pdf's of the draw successes.  Data parameter is transposed and ensured to be a dataframe
pdf_plot(as.data.frame(t(agg_data_normed)), N)
```

We can also plot the distribution of the sample for a specified number of true successes in the population
```{r, message=FALSE, warning=FALSE}
##  Plot entire sample distribution for x successes in N. Choose x (can choose more than one to compare)
x=c(25, 30)
##  first normalize by sample to get pmf and add successes column
samples_normed <- as.data.frame(data.Normalization(aggregated_data, type = "n10", normalization = "col"))
samples_normed['count'] = c(0:(nrow(samples_normed)-1))
##  plot the distribution over successes in draws for specified successes in N
draw_dist_plot(x, samples_normed, N)
```

